\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-10 paper/paper.Rnw
renamed Rnw, working on code

\SweaveOpts{concordance=TRUE}

@@

\newpage
\spacingset{1.45} % DON'T change the spacing!
\newpage
<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(splitstackshape)
correl <- read.csv("correlations.csv")
full <- read.csv("full.csv") 
Turk <- read.csv("turk.csv")

@@

\begin{figure}[htbp]
  \centering
<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(123)
uns <- gen_true_data(96, 0.8)
uns_l <- melt(uns, id="t")
p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

set.seed(123)
sm <- gen_true_data(96, 0.8, smoothed = TRUE)
sm_l <- melt(sm, id="t")
p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

@@

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
ca <- cor(a$X1, a$X2)
cb <- cor(b$X1, b$X2)
a_l <- melt(a, id="t")
b_l <- melt(b, id="t")
p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1, p2, p3, p4, ncol=2)
@
  \label{designs}
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(title="Effect of sample size, with unsmoothed data", y="Percent Correct", x="Correlation") + theme(legend.position="bottom")}
@
 \label{samplesize}
 \caption{Effects of increasing sample size for unsmoothed data. Both plot designs appear to benefit from increased sample sizes}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<confidence2, dependson='confidence', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)
y <- data.frame(table(full$subj_id))
keep2 <- y$Var1[y$Freq>9]
sub <- filter(full, subj_id %in% keep2)
sub <- sub[c(7, 15, 16, 17, 20, 27)]
sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
sub <- mutate(sub, prop = correct/Freq)
ggplot(data=sub, aes(x=conf, y= prop)) + geom_point() + labs(title="Accuracy versus Confidence", y = "Accuracy", x = "Average Confidence")
@
  \label{confidence2}
  \caption{Individual subject accuracy versus confidence. There does not appear to be any relationship between the two.}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
correl <- read.csv("correlations.csv")
pin <- full[c(5, 6, 10, 15, 16, 17)]
pin <- cSplit(pin, "response_0", direction="long")
counts <- data.frame(table(pin$pic_id, pin$response_0))
counts$Var1 <- as.integer(counts$Var1)
counts$Var2 <- as.integer(counts$Var2)
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"), smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=subset(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")

@@

Table 1: Mixed Effects Logit Model for the predicted power of plot design. Sign = 1 if correlation > 0, 0 otherwise. Each plot design is estimated separately. Results must be transformed via the logit link function to generate a power prediction. Subjects with less than ten lineup evaluations were removed from the sample. Sample size for scatter plot model: 841, overlaid lines model: 843. The model is fitted using the lme4 package in 
<<powerfit, dependson='setup', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

@@

m8l <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=subset(full.sub, plot == "L"), family="binomial")
m8s <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=subset(full.sub, plot == "S"), family="binomial")
@
%Code to make a table to be added

@@

Table 2: Mixed Effects Linear Model with the dependent variable log(time). Sign = 1 if correlation > 0, 0 otherwise. Both plot designs estimated together with Overlaid Lines as the base design. Sample size: 1684. The model is fitted using the lme4 package in R.
<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + test_param + test_param*smoothed 
                  + test_param*r + test_param*n + test_param*rsgn*r + test_param*correct + (1|subj_id), data=full.sub)
@
%get table code

>>>>

@@

\newpage
\spacingset{1.45} % DON'T change the spacing!
\newpage
<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(reshape2)
# Convert reshape2 to tidyr, melt to gather
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(splitstackshape)
library(readr)
library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv")
full <- read.csv("full.csv", row.names=1) 
Turk <- read.csv("turk.csv")

@@

\begin{figure}[htbp]
  \centering
<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(123)
uns <- gen_true_data(96, 0.8)
uns_l <- melt(uns, id="t")
p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

set.seed(123)
sm <- gen_true_data(96, 0.8, smoothed = TRUE)
sm_l <- melt(sm, id="t")
p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

@@

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
ca <- cor(a$X1, a$X2)
cb <- cor(b$X1, b$X2)
a_l <- melt(a, id="t")
b_l <- melt(b, id="t")
p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1, p2, p3, p4, ncol=2)
@
  \label{designs}
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(title="Effect of sample size, with unsmoothed data", y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
@
 \label{samplesize}
 \caption{Effects of increasing sample size for unsmoothed data. Both plot designs appear to benefit from increased sample sizes}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<confidence2, dependson='confidence', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)
y <- data.frame(table(full$subj_id))
keep2 <- y$Var1[y$Freq>9]
sub <- filter(full, subj_id %in% keep2)
sub <- select(sub, conf_level, correlation, test_param, smoothed, n, correct, sign, plot, subj_id) #c(7, 15, 16, 17, 20, 27)]
#sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
#sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
#sub <- mutate(sub, prop = correct/Freq)
qplot(correlation, 6-conf_level, data=sub, alpha=I(0.1)) + 
  geom_smooth(method="loess") + 
  facet_grid(n~plot+smoothed) +
  labs(x = "Correlation", y = "Confidence (5=High, 1=Low)")
@
  \label{confidence2}
  \caption{Individual subject accuracy versus confidence. There does not appear to be any relationship between the two.}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
# correl <- read.csv("correlations.csv") # You have already read this
pin <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
pin <- cSplit(pin, "response_0", direction="long")
counts <- data.frame(table(pin$pic_id, pin$response_0))
counts$Var1 <- as.integer(counts$Var1)
counts$Var2 <- as.integer(counts$Var2)
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
#colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(test_param=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=subset(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")

@@

Table 1: Mixed Effects Logit Model for the predicted power of plot design. Sign = 1 if correlation > 0, 0 otherwise. Each plot design is estimated separately. Results must be transformed via the logit link function to generate a power prediction. Subjects with less than ten lineup evaluations were removed from the sample. Sample size for scatter plot model: 841, overlaid lines model: 843. The model is fitted using the lme4 package in

<<powerfit, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

@@

m8l <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial")
m8s <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial")
@
%Code to make a table to be added

@@

Table 2: Mixed Effects Linear Model with the dependent variable log(time). Sign = 1 if correlation > 0, 0 otherwise. Both plot designs estimated together with Overlaid Lines as the base design. Sample size: 1684. The model is fitted using the lme4 package in R.

<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + test_param + test_param*smoothed 
                  + test_param*r + test_param*n + test_param*rsgn*r + test_param*correct + (1|subj_id), data=full.sub)
@
%get table code
